---
title: "Expedia Training Data-set Analysis"
author: "SarahLynne Palomo"
date: "26/04/2021"
output: pdf_document
highlight: tango
---

## Exploring **srch_id**

Find the unique srch_id's and count how many of them are in the submission_sample data-set:
```{r}
# Load Training data-set
tr <- read.csv("training_set_VU_DM.csv")

# Find unique search id's
uniq_srch <- unique(tr$srch_id)
num_uniq_ids <- nrow(as.matrix(uniq_srch))
head(uniq_srch, 10)
tail(uniq_srch, 10)

last_srch_id <- tail(uniq_srch, n=1)  # Highest srch_id in ascending list
perc_missing_srch <- (1 - (num_uniq_ids/last_srch_id)) * 100
perc_missing_srch
```
Here we can see that a good portion of srch_id's are missing from the training data-set sequence (40%)

## Exploring **NULL values**

```{r}
#which(is.null(tr$srch_id))  # <-- this does not work!
#which(tr$srch_id == "NULL")  # Not a real NULL value; it's only a string
#which(tr$comp3_rate == "NULL") # testing function on a column that contains "NULL"

# Create a new table for analyzing NULL stats
df_null_ratios <- data.frame(matrix(ncol = 3, nrow = 54))
names(df_null_ratios) <- list("Column_Name", "Number_of_Nulls", "Percentage_of_Rows")

tr_row <- nrow(tr)  
tr_col <- ncol(tr)
i <- 0

for (i in 1:tr_col) {
  # Populate column names from the training set into new table
  df_null_ratios[i, 1] <- names(tr)[i]      
  
  # Find the rows in the training set that contain string "NULL"
  null_rows <- which(tr[,i] == "NULL")      

  # Populate "NULL" counts into new table
  num_nulls <- nrow(as.matrix(null_rows))
  df_null_ratios[i, 2] <- num_nulls         
  
  # Populate percentage of counts into new table  
  null_perc <- (num_nulls / tr_row) * 100
  df_null_ratios[i, 3] <- null_perc         
}

df_null_ratios

```
Check the fields with more than 90% NULL values against the Assignment description of the data-set for significance!



## Exploring **Date/Time**

Dates are not in sequential order. It may be necessary to split *date_time* field into Date and Time.
```{r}
min(tr$date_time)
max(tr$date_time)
```

