---
title: "Expedia Training Data-set Analysis"
author: "SarahLynne Palomo"
date: "29/04/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
highlight: tango
---



## Exploring **NULL** values

```{r}
# Load Training data-set
tr <- read.csv("training_set_VU_DM.csv")

#which(is.null(tr$srch_id))  # <-- this does not work!
#which(tr$srch_id == "NULL")  # Not a real NULL value; it's a string
#which(tr$comp3_rate == "NULL") # testing function on a column that contains "NULL"

# Create a new table for analyzing NULL stats
df_null_ratios <- data.frame(matrix(ncol = 3, nrow = 54))
names(df_null_ratios) <- list("Column_Name", "NULL_Count", "NULL_Percentage")

tr_row <- nrow(tr)  
tr_col <- ncol(tr)
i <- 0

for (i in 1:tr_col) {
  # Populate column names from the training set into new table
  df_null_ratios[i, 1] <- names(tr)[i]      
  
  # Find the rows in the training set that contain string "NULL"
  null_rows <- which(tr[,i] == "NULL")      

  # Populate "NULL" counts into new table
  num_nulls <- nrow(as.matrix(null_rows))
  df_null_ratios[i, 2] <- num_nulls         
  
  # Populate percentage of counts into new table  
  null_perc <- (num_nulls / tr_row) * 100
  df_null_ratios[i, 3] <- null_perc         
}

df_null_ratios

```
Check the fields with more than 90% NULL values against the following descriptions of the data-set for significance!
(e.g. Not all NULLs can be converted to 0 since these would become actual values.)


---
# Data-set Descriptions


---
## Exploring **srch_id**
Data Type: integer
Description: The ID of the search

```{r}
# Find unique search id's and count how many of them are in the submission_sample data-set
uniq_srch <- unique(tr$srch_id)
num_uniq_srch <- nrow(as.matrix(uniq_srch))
head(uniq_srch, 10)
tail(uniq_srch, 10)

last_srch_id <- tail(uniq_srch, n=1)  # Highest srch_id in (incidentally) ascending list
perc_missing_srch <- (1 - (num_uniq_srch / last_srch_id)) * 100
perc_missing_srch

```
Observations: Here, we can see that a good portion of *srch_id* values are missing from the training data-set sequence (40%)


---
## Exploring **date_time**
Data Type: Date/time
Description: Date and time of the search

```{r}
min(tr$date_time)
max(tr$date_time)

```
Observations: Dates are not in sequential order. It may be necessary to split *date_time* field into Date and Time fields.


---
## Exploring **site_id**
Data Type: Integer
Description: ID of the Expedia point of sale (i.e. Expedia.com, Expedia.co.uk, Expedia.co.jp, ..)

```{r}
uniq_site <- unique(tr$site_id)
sort(uniq_site)

```
Observations: sequence is complete! Bookings can be made from multiple countries. *site_id* is specific to *srch_id*


---
## Exploring **visitor_location_country_id**
Data Type: Integer
Description: The ID of the country the customer is located

```{r}
uniq_visitor_country <- unique(tr$visitor_location_country_id)
sort(uniq_visitor_country)
num_uniq_v_country <- nrow(as.matrix(uniq_visitor_country))
max_uniq_v_country <- max(uniq_visitor_country)
missing_v_countries_perc <- (max_uniq_v_country - num_uniq_v_country) / max_uniq_v_country * 100
missing_v_countries_perc

```
Observations: 9% of visitor countries are missing from the data-set. TODO: check whether *visitor_location_country_id* is specific to *srch_id*


---
## Exploring **visitor_hist_starrating**
Data Type: Float
Description: The mean star rating of hotels the customer has previously purchased; null signifies there is no purchase history on the customer

```{r}
# Find the range of mean hotel rating scores of the non-NULL set
vhs <- subset(tr$visitor_hist_starrating, tr$visitor_hist_starrating != "NULL")
vhs <- as.numeric(vhs)
range(vhs)
hist(vhs, xlab = "visitor_hist_starrating", main = "Histogram of mean hotel rating scores")

```
Observations: Note that the mean hotel rating (*visitor_hist_starrating*) is specific to an individual (*srch_id*)


---
## Exploring **visitor_hist_adr_usd**
Data Type: Float
Description: The mean price per night (in US$) of the hotels the customer has previously purchased; null signifies there is no purchase history on the customer

```{r}
#Find the range of mean USD price per night of the non-NULL set
vha_usd <- subset(tr$visitor_hist_adr_usd, tr$visitor_hist_adr_usd != "NULL")
vha_usd <- as.numeric(vha_usd)
range(vha_usd)
hist(vha_usd, xlab = "visitor_hist_adr_usd", main = "Histogram of mean USD paid per night")

```
Observations: Most common mean price per night is 200 USD.


---
## Exploring **prop_country_id**
Data Type: Integer
Description: The ID of the country the hotel is located in

```{r}
uniq_prop_country <- unique(tr$prop_country_id)
sort(uniq_prop_country)
nrow(as.matrix(uniq_prop_country)) # 230
(230-172) / 230 * 100

```
Observations: 58/230 of property countries are missing from the data-set sequence (25%). *prop_country_id* is specific to *prop_id*


---
## Exploring **prop_id**
Data Type: Integer
Description: The ID of the hotel

```{r}
uniq_prop_id <- unique(tr$prop_id)
#sort(uniq_prop_id)
nrow(as.matrix(uniq_prop_id))  # 129113
range(uniq_prop_id)  # 1 to 140821
(140821 - 129113) / 140821 * 100

```
Observations: 11708/140821 of property id's are missing from from the data-set sequence (8.3%). 


---
## Exploring **prop_starrating**
Data Type: Integer
Description: The star rating of the hotel, from 1 to 5, in increments of 1. A 0 indicates the property has no stars, the star rating is not known or cannot be publicized

```{r}
nrow(as.matrix(tr$prop_starrating))
hist(tr$prop_starrating, main = "Histogram of hotel star ratings")

prop_nostars <- subset(tr$prop_starrating, tr$prop_starrating == 0)
nrow(as.matrix(prop_nostars))
169572 / 4958347 * 100

```

Observations: 
Property star rating (*prop_starrating*) is specific to each property listing (*prop_id*). 
Most properties are rated 3 stars.
169572/4958347 of property listings have 0 stars (3.4%)


---
## Exploring **prop_review_score**
Data Type: Float
Description: The mean customer review score for the hotel on a scale out of 5, rounded to 0.5 increments. A 0 means there have been no reviews, null that the information is not available

```{r}
#subset(tr$prop_review_score, tr$prop_id == 10404)  # example hotel
hist(tr$prop_review_score, main = "Histogram of accommodation mean review score")
subset(tr, is.null(tr$prop_review_score))

```
Observations: 
*prop_review_score* is specific to *prop_id*. 
Majority rating is 4.0 closely followed by 4.5. 
No null values found.


---
## Exploring **prop_brand_bool**
Data Type: Integer
Description: +1 if the hotel is part of a major hotel chain; 0 if it is an independent hotel

```{r}
hist(tr$prop_brand_bool, main = "Major hotel chain (1) vs independent hotel (0)")
chain_hotel <- subset(tr$prop_brand_bool, tr$prop_brand_bool == 1)
nrow(as.matrix(chain_hotel))  # 3147060
ind_hotel <- subset(tr$prop_brand_bool, tr$prop_brand_bool == 0)
nrow(as.matrix(ind_hotel))  # 1811287

nrow(as.matrix(tr$prop_brand_bool))
3147060 / 4958347 * 100
1811287 / 4958347 * 100

```

Observations: Hotel chains make up the majority of accommodation listings (63.5%) compared to independent hotels (36.5%)


---
## Exploring **prop_location_score1**
Data Type: Float
Description: A (first) score outlining the desirability of a hotel’s location

```{r}
range(tr$prop_location_score1)
hist(tr$prop_location_score1, main = "Histogram of hotel location's desirability score (1st)")

```

Observations: First score only goes up to 6.98 (out of what total??). 
Questions: What is better: a high or low score? Is this score visible to the customer while booking on the website?


---
## Exploring **prop_location_score2**
Data Type: Float
Description: A (second) score outlining the desirability of the hotel’s location

```{r}
pls2 <- subset(tr$prop_location_score2, tr$prop_location_score2 != "NULL")
pls2 <- as.numeric(pls2)
range(pls2)
hist(pls2, xlab = "prop_location_score2", main = "Histogram of hotel location's desirability score (2nd)")

```

Observations: Second score data appear to be logarithmic. 22% NULL values
Questions: Why have a second score? What is it for compared to the first score?


---
## Exploring **prop_log_historical_price**
Data Type: Float
Description: The logarithm of the mean price of the hotel over the last trading period. A 0 will occur if the hotel was not sold in that period

```{r}
range(tr$prop_log_historical_price)
hist(tr$prop_log_historical_price, main = "Histogram of log of mean hotel price in last trading period")
```
Observations: Plot of mean hotel price has a somewhat normal form
Questions: What is the calculation for the last trading period?


---
## Exploring **price_usd**
Data Type: Float
Description: Displayed price of the hotel for the given search. Note that different countries have different conventions regarding displaying taxes and fees and the value may be per night or for the whole stay

```{r}
range(tr$price_usd)  # 0 to 19726328  <--- seriously... WTF?
which.max(tr$price_usd)  # find the row index of this *ridiculously expensive* hotel
tr[1168567,]  # data-set entry for the most expensive hotel of them all

price_hotel <- subset(tr$price_usd, tr$price_usd > 0)
price_hotel <- sort(price_hotel)
tail(price_hotel)   # looks like it wasn't just an outlier. Wow. Who books this stuff online??
head(price_hotel)
hist(tr$price_usd, main = "Histogram of hotel prices in USD")
```

Observations: 
Need to check *price_usd* against *site_id* and *srch_length_of_stay* to normalise the value to per night due to different taxes and fees per country.
There are lots of 0's and weird values that increment in the lower end of the data. We will definitely have to fix this.
Questions: Are the 0 prices for searches that didn't amount to a booking? Or is it a currency conversion error?

---
## Exploring **promotion_flag**
Data Type: Integer
Description: +1 if the hotel had a sale price promotion specifically displayed

```{r}
hist(tr$promotion_flag, main = "Histogram of promotional flag")
promo_yes <- subset(tr$promotion_flag, tr$promotion_flag == 1)
nrow(as.matrix(promo_yes))  # 1069118
promo_no <- subset(tr$promotion_flag, tr$promotion_flag == 0)
nrow(as.matrix(promo_no))  # 3889229
1069118 / 4958347 * 100  # 21.6%
3889229 / 4958347 * 100  # 78.4%

```
Observations: 21.6% of accommodations are listed with a promotional sale price. This really should be type boolean instead of integer. No NULLs found.


---
## Exploring **srch_destination_id**
Data Type: Integer
Description: ID of the destination where the hotel search was performed

```{r}
uniq_dest_id <- unique(tr$srch_destination_id)
uniq_dest_id <- sort(uniq_dest_id)
range(uniq_dest_id)  #   2 to 28416
nrow(as.matrix(uniq_dest_id))  # 18127
(28416 - 18127) / 28416 * 100  # 36.20847
hist(tr$srch_destination_id, main = "Histogram of destination IDs")

```
Observations: Some destinations are more popular than others (TODO: find out what ones are these). 10289/28416 of destination ids are missing from the data-set sequence (36%)


---
## Exploring **srch_length_of_stay**
Data Type: Integer
Description: Number of nights stay that was searched

```{r}
hist(tr$srch_length_of_stay, main = "Histogram number of nights searched (all)")
range(tr$srch_length_of_stay)

num_nights10 <- subset(tr$srch_length_of_stay, tr$srch_length_of_stay <= 10)
hist(num_nights10, main = "Histogram number of nights searched (<= 10)")

```
Observations: Data dsitribution appears logarithmic. Most common stay duration is 1 night then 2 nights.

---
##Exploring **srch_booking_window**
Data Type: Integer
Decription: Number of days in the future the hotel stay started from the search date

```{r}
range(tr$srch_booking_window)
hist(tr$srch_booking_window, main = "Histogram of booking window")
days_future100 <- subset(tr$srch_booking_window, tr$srch_booking_window <= 100)
hist(days_future100, main = "Histogram of booking window (100 days)")
```
Observations: Data dsitribution appears logarithmic. Most common booking window is within 5 days of stay


---
## Exploring **srch_adults_count**
Data Type: Integer
Description: The number of adults specified in the hotel room

```{r}
range(tr$srch_adults_count)
hist(tr$srch_adults_count, main = "Histogram of number of adults per booking")

```
Observations: Most common hotel booking is for 2 adults per room.


---
## Exploring **srch_children_count**
Data Type: Integer
Description: The number of (extra occupancy) children specified in the hotel room

```{r}
range(tr$srch_children_count)
hist(tr$srch_children_count, main = "Histogram of number of children per booking")

```
Observations: Most bookings are made without including children


---
## Exploring **srch_room_count**
Data Type: Integer
Description: Number of hotel rooms specified in the search

```{r}
range(tr$srch_room_count)
hist(tr$srch_room_count, main = "Histogram of number of rooms")

```
Observations: Vast majority of bookings are made for 1 room only


---
## Exploring **srch_saturday_night_bool**
Data Type: Boolean
Description: +1 if the stay includes a Saturday night, starts from Thursday with a length of stay is less than or equal to 4 nights (i.e. weekend); otherwise 0

```{r}
hist(tr$srch_saturday_night_bool, main = "Histogram of Saturday night boolean")
saturday_no <- subset(tr$srch_saturday_night_bool, tr$srch_saturday_night_bool == 0)
nrow(as.matrix(saturday_no))  # 2468202
saturday_yes <- subset(tr$srch_saturday_night_bool, tr$srch_saturday_night_bool == 1)
nrow(as.matrix(saturday_yes))  # 2490145

```

Observations: Number of bookings made including Saturday night (2490145) is almost the same as excluding (2468202)


---
## Exploring **srch_query_affinity_score**
Data Type: Float
Description: The log of the probability a hotel will be clicked on in Internet searches (hence the values are negative) A null signifies there are no data (i.e. hotel did not register in any searches)

```{r}
#range(tr$srch_query_affinity_score, na.rm = FALSE)  # excluding NULL  <-- this doesn't work properly

sqas <- subset(tr$srch_query_affinity_score, tr$srch_query_affinity_score != "NULL")
sqas <- as.numeric(sqas)
range(sqas)
hist(sqas, xlab = "srch_query_affinity_score", main = "Histogram of search query affinity score")

```

Observations: 93.6% of the data values are NULL. Even though the aata is officially type float, it was actually type char so it first had to be converted to type numeric for the histogram and to give the correct range. Histogram excludes NULL values.


---
## Exploring **orig_destination_distance**
Data Type: Float
Description: Physical distance between the hotel and the customer at the time of search. A null means the distance could not be calculated

```{r}
#range(tr$orig_destination_distance)
o2d_dist <- subset(tr$orig_destination_distance, tr$orig_destination_distance != "NULL")
o2d_dist <- as.numeric(o2d_dist)
range(o2d_dist)
hist(o2d_dist)

```

Observations: 32.4% of the data values are NULL. Converted type char to numeric. Histogram excludes NULL values. It's not clear in what units the distance is calculated in. Most bookings are made when the customer is close to the hotel


---
## Exploring **random_bool**
Data Type: Boolean
Description: +1 when the displayed sort was random, 0 when the normal sort order was displayed

```{r}
hist(tr$random_bool, main = "Histogram of random boolean on display sort")
sort_random <- subset(tr$random_bool, tr$random_bool == 1)
nrow(as.matrix(sort_random))  # 1467177
sort_normal <- subset(tr$random_bool, tr$random_bool == 0)
nrow(as.matrix(sort_normal))  # 3491170
1467177 / tr_row * 100  # 29.59004
3491170 / tr_row * 100  # 70.40996
```
Observations: Random sort display = 29.6%; Normal sort display = 70.4%


---
## Exploring **comp1_rate**
Data Type: Integer
Description: +1 if Expedia has a lower price than competitor 1 for the hotel; 0 if the same; -1 if Expedia’s price is higher than competitor 1; null signifies there is no competitive data

```{r}
c1_rate <- subset(tr$comp1_rate, tr$comp1_rate != "NULL")
c1_rate <- as.numeric(c1_rate)
hist(c1_rate)

c1_rate_lower <- subset(c1_rate, c1_rate == 1)
c1_rate_same <- subset(c1_rate, c1_rate == 0)
c1_rate_higher <- subset(c1_rate, c1_rate == -1)

nrow(as.matrix(c1_rate_lower))  # 67256
nrow(as.matrix(c1_rate_same))  # 42959
nrow(as.matrix(c1_rate_higher))  # 9715

nrow(as.matrix(c1_rate))  # 119930

67256 / 119930 * 100  # 56.07938
42959 / 119930 * 100  # 35.82006
9715 / 119930 * 100   # 8.100559

```

Observations: Data contains 97.6% NULL values but we cannot change these to 0 as NULL has a meaning. Of the remaining data - 56.1% Expedia has a lower price; 35.8% Expedia's price is the same as competitor; 8.1% Expedia has a higher price.


---
## Exploring **comp1_inv**
Data Type: +1 if competitor 1 does not have availability in the hotel; 0 if both Expedia and competitor 1 have availability; null signifies there is no competitive data

```{r}

```

Observations: 97.4% of values are NULL